### 2016-11-25 amenIntro/amenIntro.Rmd
push up most of intro, waiting on mod results

---
title: "Networks Matter...or some nonsense like that"
output: 
      ioslides_presentation:
            smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Purpose

* Why are we meeting during Thanksgiving weekend...I didn't realize the date and none of you objected :)
* Anyhow...we are here to talk about conducting a meta-analysis of dyadic studies in political science using the Additive and Multiplicative Effects Model for Networks (AMEN)
* What I hope we find from this project is that using standard GLM approaches on dyadic data will produce
      - biased estimates of the effect of independent variables
      - uncalibrated confidence intervals
      - and poor predictive performance.
* Anticipated outputs: 
      - One big meta-analysis piece in which we show the items in the list above to be true on average
      - A number of small replication pieces

## Outline

* **Brief** background discussion of AMEN based on [Hoff's vignette](https://arxiv.org/pdf/1506.08237v1.pdf) and a [joint piece with me and Mike](https://arxiv.org/pdf/1611.00460v1.pdf)
* Example application of AMEN, where we will   
      - Discuss how to structure input data
      - Extract and interpret model results
      - Conduct out-of-sample performance on longitudinal networks
* Discuss project management

## Dyadic data

* Any dataset in which the unit of observation is characterized by a pair of actors
      - MIDs
      - International economic agreements (PTAs, BITs)
      - Alliances (NATO)
      - Trade flows

## What makes dyadic data special

* Formation of a tie between any pair of actors is **not necessarily** independent of other ties
      - This is true for bilateral and multilateral events
* Why?

## Actors are part of a system ... a network

```{r, out.width= "700px",fig.align="center"}
knitr::include_graphics("dyaddata.png")
```

## Flavors of dependencies

* First order dependence
      - Often will find significant heterogeneity in activity levels across nodes
* Second order dependence
      - Reciprocity: Event sent from i $\rightarrow$ j is likely related to an event sent from j $\rightarrow$ i

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \color{red}{\mu} + \color{red}{e_{ij}} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\mu$ baseline measure of network activity, for the purpose of regression we turn this into $\beta^{T}X$
* $e_{ij}$ residual variation that we will use the SRM to decompose

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= \color{red}{a_{i} + b_{j}} + \epsilon_{ij} \\
      \color{red}{\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \}} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* row/sender effect ($a_{i}$) & column/receiver effect ($b_{j}$)
* Modeled jointly to account for correlation in how active an actor is in sending and receiving ties

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\color{red}{\Sigma_{ab}}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \color{red}{\Sigma_{ab}} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$ capture heterogeneity in the row and column means
* $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties)

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \color{red}{\epsilon_{ij}} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \color{red}{\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\}} &\sim N(0,\color{red}{\Sigma_{\epsilon}}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; & \color{red}{\Sigma_{\epsilon}} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\epsilon_{ij}$ captures the within dyad effect
* Second-order dependencies are described by $\sigma_{\epsilon}^{2}$
* Within dyad correlation, aka reciprocity, represented by $\rho$

## Network level dependencies

```{r, out.width= "700px",fig.align="center"}
knitr::include_graphics("homophStoch.png")
```

* Homophily
* Stochastic equivalence

## Latent variable models for networks

$$
\begin{aligned}
\text{Latent class model} \\
      &\alpha(u_{i}, u_{j}) = m_{u_{i}, u_{j}} \\
      &u_{i} \in \{1, \ldots, K \}, \; i \in \{1,\ldots, n\} \\
      &M \text{ a } K \times K \text{ symmetric matrix} \\
\text{Latent distance model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = -|\textbf{u}_{i} - \textbf{u}_{j}| \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
\text{Latent factor model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{T} \Lambda \textbf{u}_{j} \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
      &\Lambda \text{ a } K \times K \text{ diagonal matrix}
\end{aligned}
$$

## Our preferred approach

$$
\begin{aligned}
\text{Latent class model} \\
      &\alpha(u_{i}, u_{j}) = m_{u_{i}, u_{j}} \\
      &u_{i} \in \{1, \ldots, K \}, \; i \in \{1,\ldots, n\} \\
      &M \text{ a } K \times K \text{ symmetric matrix} \\
\text{Latent distance model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = -|\textbf{u}_{i} - \textbf{u}_{j}| \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
\color{red}{\text{Latent factor model}} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{T} \Lambda \textbf{u}_{j} \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
      &\Lambda \text{ a } K \times K \text{ diagonal matrix}
\end{aligned}
$$

## Putting it all together: AME

$$
\begin{aligned}
      y_{ij} &= g(\theta_{ij}) \\ 
      &\theta_{ij} = \beta^{T} \mathbf{X}_{ij} + e_{ij} \\
      &e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \textbf{u}_{i}^{T} \textbf{D} \textbf{v}_{j} \\
\end{aligned}
$$

* $a_{i} + b_{j}  + \epsilon_{ij}$, are additive random effects and account for sender, receiver, and within-dyad dependence
* multiplicative effects, $\textbf{u}_{i}^{T} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns that are left over in $\theta$ after accounting for any known covariate information

## Benefits of this approach

* At its core, AME is just a GLM with random effects used to ensure that we can treat dyadic observations as conditionally independent
* AME can be used: 
      - on both undirected and directed data, 
      - on longitudinal and static networks,
      - and on a variety of distribution types we commonly encounter in political science (binomial, gaussian, and ordinal).

## How to use it...first installing

```{r,echo=TRUE, message=FALSE, warning=FALSE, suppressPackageStartupMessages=TRUE}
library(devtools)
devtools::install_github('s7minhas/amen')
library(amen)
```

**DON'T USE THE VERSION FROM CRAN**

## Structuring input data

* This is going to be a bit different than what is described in [Hoff's vignette](https://arxiv.org/pdf/1506.08237v1.pdf). 
      - As of this week, AMEN can now be estimated on networks where the actor composition changes over time, and to make that change we switch to using **lists** instead of **arrays**.
* In undirected longitudinal setting three inputs:
      - `Y`: a T length list of n x n adjacency matrices (see Table 2)
      - `Xdyad`: A T length list of n x n x p arrays
      - `Xrow`: A T length list of n x p matrices
* In directed longitudinal setting you can distinguish between:
      - Sender covariates `Xrow`
      - Receiver covariates `Xcol`

## Structuring dependent variable

```{r,echo=FALSE}
load('~/Dropbox/Research/netsMatter/replications/example/inputData/exampleData.rda')
```

```{r, echo=TRUE}
str(yL)
```

## More on dependent variable

```{r, echo=TRUE}
print( yL$'1985'[17:33,17:33]  )
```

## Dyadic covariate(s)

```{r, echo=TRUE}
str(xDyadL)
```

## More on dyadic covariate(s)

```{r, echo=TRUE}
print( round(xDyadL$'1985'[17:33,17:33,1],1)  )
```

## Actor level covariate(s)

```{r, echo=TRUE}
str(xNodeL)
```

## More on actor level covariate(s)

```{r, echo=TRUE}
print( xNodeL$'1985'[17:33,]  )
```

## Running an AME model

```{r, out.width= "600px",fig.align="center"}
knitr::include_graphics("ameRepL.png")
```

## Lets put it all together

* Undirected, `symmetric=TRUE`, binary, `model='bin'`, longitudinal AME model with actor random effects, `nvar=TRUE`, and a two dimensional multiplicative effect, `R=2`.

```{r, echo=TRUE,eval=FALSE}
fit=ame_repL(Y=yL, Xdyad=xDyadL, Xrow=xNodeL, Xcol=NULL,
      nvar=TRUE, R=2,
      model="bin",symmetric=TRUE,
      burn=500,nscan=1000,odens=10, # WILL NEED TO BE LARGER
      plot=FALSE, print=FALSE, 
      seed=6886 # ALWAYS SET A SEED
      )
```

```{r, echo=TRUE}
load('~/Dropbox/Research/netsMatter/replications/example/outputData/model_k2.rda')
```

## The output

```{r, out.width= "600px",fig.align="center"}
knitr::include_graphics("ameOut.png")
```

## Assessing convergence

## Did you set the right K?

## Exogenous covariates

```{r, echo=TRUE}
dim(fit$BETA)

stats = function(x){ c( mu=mean(x), q95  ) }
apply(BETA<)
```

## Out of sample model performance