### 2018-03-01 paper/netsMatter.tex
abstract edits

\documentclass[12pt]{amsart} 
      % \newif\ifpreview
      % \previewfalse

% approx iid
\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}
\usepackage{natbib}
\usepackage{titlecaps}
\Addlcwords{a about above after again against all am an and any are aren't as at be because been before being below between both but by can't cannot could couldn't did didn't do does doesn't doing don't down during each few for from further had hadn't has hasn't have haven't having he he'd he'll he's her here here's hers herself him himself his how how's i i'd i'll i'm i've if in into is isn't it it's its itself let's me more most mustn't my myself no nor not of off on once only or other ought our ours ourselves out over own same shan't she she'd she'll she's should shouldn't so some such than that that's the their theirs them themselves then there there's these they they'd they'll they're they've this those through to too under until up very was wasn't we we'd we'll we're we've were weren't what what's when when's where where's which while who who's whom why why's with won't would wouldn't you you'd you'll you're you've your yours yourself yourselves}

% Fonts
% at some point figure out bolding ...
\usepackage[default,osfigures,scale=0.95]{opensans}
%\usepackage[sfdefault,scaled=.85]{FiraSans}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[varqu,varl]{zi4}% inconsolata typewriter
\usepackage{amsmath,amsthm}
\usepackage[cmintegrals]{newtxsf}
\usepackage[T1]{fontenc}
\usepackage{ae}

% easy commands for number propers
\newcommand{\first}{1$^{\text{st}}$}
\newcommand{\second}{2$^{\text{nd}}$}
\newcommand{\third}{3$^{\text{rd}}$}
\newcommand{\nth}[1]{${#1}^{\text{th}}$}

% Generate some fake text
\usepackage{blindtext}

% % add watermark
% \usepackage{draftwatermark}
% \SetWatermarkText{Version 1.0}
% \SetWatermarkScale{.5}
% \SetWatermarkColor[gray]{0.9}

% layout control
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
\usepackage{parallel}
\usepackage{parcolumns}
\usepackage{fancyhdr}
\usepackage[export]{adjustbox}
\usepackage{etoolbox}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{relsize}
\usepackage{mathtools}
\usepackage{bm}
%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}

% restricts float objects to be inserted before end of section
% creates float barriers
\usepackage[section]{placeins}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{longtable}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% footnotes at bottom
%\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[pdftex,hyperref,x11names]{xcolor}

% for multiple references and insertion of the word "figure" or "table"
% \usepackage{cleveref}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
\usepackage{tikz}

% sideway figures
\usepackage{rotating}
\usepackage{lscape}

%\usepackage[figuresright]{rotating}
% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
%   {\end{minipage}\end{sidewaysfigure}}

% Spacing
\usepackage[doublespacing]{setspace}

% outside code
\usepackage{listings}
\usepackage{enumitem}

% define clickable links and their colors
\hypersetup{%
	unicode=false,          % non-Latin characters in Acrobat's bookmarks
	pdftoolbar=true,        % show Acrobat's toolbar?
	pdfmenubar=true,        % show Acrobat's menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfnewwindow=true,%
	%pagebackref=false,%
	pdfauthor={Shahryar Minhas, et alia.},%
	pdftitle={Networks Matter},%
	colorlinks,%
	citecolor=black,%
	filecolor=black,%
	linkcolor=black,%
	urlcolor=RoyalBlue4
}

% -------------------- title -------------------- %

% \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
% \title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
\title{Taking Dyads Seriously}
\vspace{\baselineskip}

% \author[Minhas]{Shahryar Minhas}
% \address{Shahryar Minhas: Department of Political Science}
% \curraddr{Michigan State University}
% \email[Corresponding author]{minhassh@msu.edu}

% \author[Dorff]{Cassy L. Dorff}
% \address{Cassy L. Dorff}
% \curraddr{Department of Political Science, University of New Mexico}
% \email{cdorff@unm.edu}

% \author[Foster]{Margaret Foster}
% \address{Margaret Foster}
% \curraddr{Department of Political Science, Duke University}
% \email{margaret.foster@duke.edu}

% \author[Gallop]{Max Gallop}
% \address{Max Gallop}
% \curraddr{Department of Government and Public Policy, University of Strathclyde}
% \email{max.gallop@strath.ac.uk}

% \author[Liu]{Howard Liu}
% \address{Howard Liu}
% \curraddr{Department of Political Science, Duke University}
% \email{hao.liu@duke.edu}

% \author[Tellez]{Juan Tellez}
% \address{Juan Tellez}
% \curraddr{Department of Political Science, Duke University}
% \email{juan.tellez@duke.edu}

% \author[Ward]{Michael D. Ward}
% \address{Michael D. Ward}
% \curraddr{Department of Political Science, Duke University}
% \email{michael.d.ward@duke.edu}

% % \date{\today}

% \thanks{Shahryar Minhas and Michael D. Ward acknowledge support from National Science Foundation (NSF) Award 1259266. John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments. We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
% \chead{\fancyplain{}{Taking Dyads Seriously}}
\chead{\fancyplain{}{}}
% \rhead{\fancyplain{}{\today}}
\rhead{\fancyplain{}{}}
\rfoot{\fancyplain{}{\thepage}}

% references to graphics
\makeatletter
% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}

% easy command for boldface math symbols
\newcommand{\mbs}[1]{\boldsymbol{#1}}

% command for R package font
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

% Add some colors
\definecolor{red1}{RGB}{253,219,199}
\definecolor{red2}{RGB}{244,165,130}
\definecolor{red3}{RGB}{178,24,43}

\definecolor{green1}{RGB}{229,245,224}
\definecolor{green2}{RGB}{161,217,155}
\definecolor{green3}{RGB}{49,163,84}

\definecolor{blue0}{RGB}{255,247,251}
\definecolor{blue1}{RGB}{222,235,247}
\definecolor{blue2}{RGB}{158,202,225}
\definecolor{blue3}{RGB}{49,130,189}
\definecolor{blue4}{RGB}{4,90,141}

\definecolor{purple1}{RGB}{191,211,230}
\definecolor{purple2}{RGB}{140,150,198}
\definecolor{purple3}{RGB}{140,107,177}

\definecolor{brown1}{RGB}{246,232,195}
\definecolor{brown2}{RGB}{223,194,125}
\definecolor{brown3}{RGB}{191,129,45}

% square bracket matrices
\let\bbordermatrix\bordermatrix
\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}

\begin{document}
% \ifpreview {
%    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
%    \pagecolor{personal}
%    }
% \fi

\small{\singlespacing{
\begin{abstract}
	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data. Yet, contemporary research brushes aside the complications that come with analyzing interdependent relational data. Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches. As a result, studies often produce results with biased coefficient estimates and poorly calibrated standard errors. These biases have profound consequences for evaluating parametric models. We present an alternative, regression-based approach that accounts for the dependencies complicating this type of analysis. We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data. In addition, we replicate three recent studies in recent international relations scholarship, and compare  the standard approach to our alternative. For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables. Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and for  modeling the data generating process behind outcomes of interest.
\end{abstract}
}}

\maketitle\thispagestyle{empty}
\newpage\setcounter{page}{1}

\input{1_intro}
\input{2_model}
\input{3_sim}
\input{4_app}
\input{5_concl}
\input{6_appendix}

\newpage
% Bib stuff
\clearpage

%\bibliography{/Users/maxgallop/documents/whistle/master}
\bibliography{/Users/s7m/whistle/master}
 % \bibliography{/Users/mdw/Git/whistle/master}
% \bibliography{/Users/cassydorff/ProjectsGit/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}

\end{document}