### 2018-03-02 paper/netsMatter.tex
abstract

\documentclass[12pt]{amsart} 
      % \newif\ifpreview
      % \previewfalse

% approx iid
\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}
\usepackage{natbib}
\usepackage{titlecaps}
\Addlcwords{a about above after again against all am an and any are aren't as at be because been before being below between both but by can't cannot could couldn't did didn't do does doesn't doing don't down during each few for from further had hadn't has hasn't have haven't having he he'd he'll he's her here here's hers herself him himself his how how's i i'd i'll i'm i've if in into is isn't it it's its itself let's me more most mustn't my myself no nor not of off on once only or other ought our ours ourselves out over own same shan't she she'd she'll she's should shouldn't so some such than that that's the their theirs them themselves then there there's these they they'd they'll they're they've this those through to too under until up very was wasn't we we'd we'll we're we've were weren't what what's when when's where where's which while who who's whom why why's with won't would wouldn't you you'd you'll you're you've your yours yourself yourselves}

% Fonts
% at some point figure out bolding ...
\usepackage[default,osfigures,scale=0.95]{opensans}
%\usepackage[sfdefault,scaled=.85]{FiraSans}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[varqu,varl]{zi4}% inconsolata typewriter
\usepackage{amsmath,amsthm}
\usepackage[cmintegrals]{newtxsf}
\usepackage[T1]{fontenc}
\usepackage{ae}

% easy commands for number propers
\newcommand{\first}{1$^{\text{st}}$}
\newcommand{\second}{2$^{\text{nd}}$}
\newcommand{\third}{3$^{\text{rd}}$}
\newcommand{\nth}[1]{${#1}^{\text{th}}$}

% Generate some fake text
\usepackage{blindtext}

% % add watermark
% \usepackage{draftwatermark}
% \SetWatermarkText{Version 1.0}
% \SetWatermarkScale{.5}
% \SetWatermarkColor[gray]{0.9}

% layout control
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
\usepackage{parallel}
\usepackage{parcolumns}
\usepackage{fancyhdr}
\usepackage[export]{adjustbox}
\usepackage{etoolbox}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{relsize}
\usepackage{mathtools}
\usepackage{bm}
%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}

% restricts float objects to be inserted before end of section
% creates float barriers
\usepackage[section]{placeins}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{longtable}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% footnotes at bottom
%\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[pdftex,hyperref,x11names]{xcolor}

% for multiple references and insertion of the word "figure" or "table"
% \usepackage{cleveref}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
\usepackage{tikz}

% sideway figures
\usepackage{rotating}
\usepackage{lscape}

%\usepackage[figuresright]{rotating}
% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
%   {\end{minipage}\end{sidewaysfigure}}

% Spacing
\usepackage[doublespacing]{setspace}

% outside code
\usepackage{listings}
\usepackage{enumitem}

% define clickable links and their colors
\hypersetup{%
	unicode=false,          % non-Latin characters in Acrobat's bookmarks
	pdftoolbar=true,        % show Acrobat's toolbar?
	pdfmenubar=true,        % show Acrobat's menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfnewwindow=true,%
	%pagebackref=false,%
	pdfauthor={Shahryar Minhas, et alia.},%
	pdftitle={Networks Matter},%
	colorlinks,%
	citecolor=black,%
	filecolor=black,%
	linkcolor=black,%
	urlcolor=RoyalBlue4
}

% -------------------- title -------------------- %

% \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
% \title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
\title{Taking Dyads Seriously}
\vspace{\baselineskip}

% \author[Minhas]{Shahryar Minhas}
% \address{Shahryar Minhas: Department of Political Science}
% \curraddr{Michigan State University}
% \email[Corresponding author]{minhassh@msu.edu}

% \author[Dorff]{Cassy L. Dorff}
% \address{Cassy L. Dorff}
% \curraddr{Department of Political Science, University of New Mexico}
% \email{cdorff@unm.edu}

% \author[Foster]{Margaret Foster}
% \address{Margaret Foster}
% \curraddr{Department of Political Science, Duke University}
% \email{margaret.foster@duke.edu}

% \author[Gallop]{Max Gallop}
% \address{Max Gallop}
% \curraddr{Department of Government and Public Policy, University of Strathclyde}
% \email{max.gallop@strath.ac.uk}

% \author[Liu]{Howard Liu}
% \address{Howard Liu}
% \curraddr{Department of Political Science, Duke University}
% \email{hao.liu@duke.edu}

% \author[Tellez]{Juan Tellez}
% \address{Juan Tellez}
% \curraddr{Department of Political Science, Duke University}
% \email{juan.tellez@duke.edu}

% \author[Ward]{Michael D. Ward}
% \address{Michael D. Ward}
% \curraddr{Department of Political Science, Duke University}
% \email{michael.d.ward@duke.edu}

% % \date{\today}

% \thanks{Shahryar Minhas and Michael D. Ward acknowledge support from National Science Foundation (NSF) Award 1259266. John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments. We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
% \chead{\fancyplain{}{Taking Dyads Seriously}}
\chead{\fancyplain{}{}}
% \rhead{\fancyplain{}{\today}}
\rhead{\fancyplain{}{}}
\rfoot{\fancyplain{}{\thepage}}

% references to graphics
\makeatletter
% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}

% easy command for boldface math symbols
\newcommand{\mbs}[1]{\boldsymbol{#1}}

% command for R package font
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

% Add some colors
\definecolor{red1}{RGB}{253,219,199}
\definecolor{red2}{RGB}{244,165,130}
\definecolor{red3}{RGB}{178,24,43}

\definecolor{green1}{RGB}{229,245,224}
\definecolor{green2}{RGB}{161,217,155}
\definecolor{green3}{RGB}{49,163,84}

\definecolor{blue0}{RGB}{255,247,251}
\definecolor{blue1}{RGB}{222,235,247}
\definecolor{blue2}{RGB}{158,202,225}
\definecolor{blue3}{RGB}{49,130,189}
\definecolor{blue4}{RGB}{4,90,141}

\definecolor{purple1}{RGB}{191,211,230}
\definecolor{purple2}{RGB}{140,150,198}
\definecolor{purple3}{RGB}{140,107,177}

\definecolor{brown1}{RGB}{246,232,195}
\definecolor{brown2}{RGB}{223,194,125}
\definecolor{brown3}{RGB}{191,129,45}

% square bracket matrices
\let\bbordermatrix\bordermatrix
\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}

\begin{document}
% \ifpreview {
%    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
%    \pagecolor{personal}
%    }
% \fi

\small{\singlespacing{
\begin{abstract}
Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data. These types of data do not typically satisfy the conditional independence criterion required for most statistical approaches. Consequently, modeling approaches that fail to adequately capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors. We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that more fully accounts for the dependencies in dyadic data. First, we conduct a simulation that reveals how the model captures dependencies that emerge in relational data and demonstrate why accounting for these processes improves inference. Second, through replication, we compare the AME model to approaches used in three studies from recent international relations scholarship. For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME. Further, these conventional methods overstate the effect of key variables and underestimate the uncertainty in these effects. In sum, our work shows the consequences of failing to seriously consider dependencies inherent to dyadic data. Perhaps most importantly, by precisely modeling the data generating process underlying political phenomena we improve scholars’ ability to conduct inferential analyses.

\end{abstract}
}}

\maketitle\thispagestyle{empty}
\newpage\setcounter{page}{1}

\input{1_intro}
\input{2_model}
\input{3_sim}
\input{4_app}
\input{5_concl}
\input{6_appendix}

\newpage
% Bib stuff
\clearpage

%\bibliography{/Users/maxgallop/documents/whistle/master}
%\bibliography{/Users/s7m/whistle/master}
 % \bibliography{/Users/mdw/Git/whistle/master}
 \bibliography{/Users/cassydorff/ProjectsGit/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}

\end{document}