### 2018-03-05 paper/1_intro.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

\section{\textbf{Introduction}}

\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).  However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid). Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases. In practice, such strategies rarely work because they do not directly address the fundamental data generating process. This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.

In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2016:arxiv}. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict against a government, the government will likely reciprocate that behavior. We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. Third-order dependencies capture relationships of transitivity, balance, and clusterability between different dyads. For example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions. The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.

We begin with a discussion of these dependencies and an introduction to the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data. Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature (published in this journal). Our comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables. Consequently, AME generates results that, at times, differ from those found in the original study as well as from the broader literature. Moreover, we demonstrate that the AME latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data widely employed in IR. Finally, we show that for each replication our network-based approach provides substantially more accurate out-of-sample predictions than the models used in the original studies.

Our AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest. Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, but at the same time account for the data generating processes behind such events of interest. Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.