### 2020-01-17 revisionMemo/rev2.tex
move repl code to my cluster, org revision memo, start to assign tasks

\section*{Reviewer 2}

\begin{enumerate}
	\item The additive and multiplicative effects (AME) model presented in the paper allows for a better treatment of first, second and third order dependencies in the data, and account for unobserved sources of bias in the data. The presentation of the AME estimator is clear, and the simulations provide clear evidence that the model outperforms naive OLS/GLM models fitted to the same data. The paper will be of great interest to the PSRM audience and is likely to have a strong impact on future research in international relations. I do, however, have a few suggestions for revisions that will hopefully strengthen the paper’s contributions.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for their comments and suggestions.
		}}
	\end{itemize}
	\item The paper shows that the AME model performs well in comparison to the “oracle” model where all sources of dependencies in the data are accounted for, which is reassuring. Yet this comparison feels incomplete. First, the naive model is clearly misspecified given the underlying data generating process. A large number of dependencies in international relations are linked to geography. What would happen if the dependencies were modeled using spatial econometric techniques?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item It would help if the author discussed in more depth the properties of the AME estimator in the presence of unobserved dependencies. One would imagine that the oracle model, where all the unobserved factors and dependencies are accounted for should always outperform a parametric or semi-parametric estimator. Relatedly, it also would help to have a more technical discussion of the properties of AME models, particularly in terms of the bias-variance trade off.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item The author could also underscore a related issue, which usually escapes empirical IR scholars: the importance of developing theories that account for the direct and indirect effects of interactions in a setting where the units are interconnected, and where the effects on one unit have the potential to impact other units in the system. A traditional approach is to throw as many observable variables as possible as controls, without much consideration on where they fit in the model. As the comparison of the AME and oracle models suggests, a careful modeling of spillovers and dependencies can result in better estimates of the underlying relationships.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item The paper emphasizes on the importance of estimating the random effects for theory building and hypothesis test. A longer elaboration and illustration of this point would be very helpful.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item I would like to see a more detailed discussion of the setup and choices made for the Monte Carlo simulations. What type of network do the simulated data model? How does the network in the simulation relate to typical networks in IR studies, and to the ones in the replicated papers.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		SM will deal with this.
		}}
	\end{itemize}
	\item The criteria for choosing the studies for replication is unclear and needs a better justification. I would have liked to see a larger set of replications particularly in areas and topics where the data generating process, the dependencies among units varies, or the relationship between the units is hierarchical, cross-classified or multilevel.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item Given the importance of promoting the use of the AME method I would encourage the author to revise the tutorial presented in Appendix C to make it more accessible to a larger audience.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
\end{enumerate}