### 2020-01-17 revisionMemo/revisionLetter.tex

\documentclass[12pt,onesided,fullpage]{amsart}

%%% PAGE DIMENSIONS
\usepackage{multirow}
\usepackage{placeins}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
%\pdfpagewidth=8.5in % for pdflatex
%\pdfpageheight=11in % for pdflatex

\usepackage{graphicx} % support the \includegraphics command and options

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array,amsmath,graphicx,psfrag,amssymb,subfigure,tabularx,booktabs}
\usepackage{mparhack}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{multicol}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{placeins}
\setcitestyle{authordate,round,semicolon,aysep={,},yysep={,}}
\bibpunct[:]{(}{)}{;}{a}{,}{,}

% \doublespacing
\def\input@path{
	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
	{C:/Users/herme/Research/netsMatter/paper/graphics/},
	{/Users/s7m/Research/netsMatter/paper/graphics/}
}

\graphicspath{
	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
	{C:/Users/herme/Research/netsMatter/paper/graphics/},
	{/Users/s7m/Research/netsMatter/paper/graphics/},
}

\begin{document}
\singlespacing

\title[PSRM-OA-2019-0126]{Ms. No. PSRM-OA-2019-0126 entitled ``Taking Dyads Seriously''}

\date{\today~~Version 0.01}
\maketitle

$\hspace{-5mm}$Dear Editors, \\ [1ex]

Thank you for the opportunity to revise and resubmit our manuscript. We believe the manuscript has benefited from the Reviewers' helpful and thoughtful comments. We have revised the manuscript, taking seriously each individual point raised by the Reviewers. Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each of the points made by the Reviewers.

We hope you agree that the manuscript has improved through this process and we are looking forward to your response.\\ [1ex]

% Sincerely, \\ [1ex]

% The Authors

\include{rev1}
\include{rev2}

\newpage\tiny
\end{document}\bye