### 2020-01-23 paper/6_appendix.tex
started response to reviewer comment on appendix c tutorial

\clearpage

\renewcommand{\thefigure}{\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{\thesection \arabic{table}}
\setcounter{table}{0}
%\renewcommand{\thesection}{A.\arabic{section}}
\setcounter{section}{0}

\appendix
\section{AME Monte Carlo Markov Chain Algorithm}

Given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:

 \begin{itemize}
  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
  \item For each $k \in K$:
  \begin{itemize}
    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$. }
  \end{itemize}
 \end{itemize}
\newpage
\section{Additional Replication Information}

For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation and the results from the AME model.  We have adopted for presentation purposes only the use of significance testing on these observational data for facile comparison with the replicated studies.  However, note that we do provide predictive heuristics for these models as well.

% Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis. %For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.

\clearpage
\subsection*{Reiter \& Stam (2003)}

Additional information for the Reiter \& Stam (2003) re-estimation.

\input{reiter_stam_coeftable}
\FloatBarrier

% \begin{figure}
% 	\centering   
% 	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
% 	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
% 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves, PR curves, and separation plots.}
% \end{figure}
\FloatBarrier
\clearpage

\subsection*{Weeks (2012)}

Additional information for the Weeks (2012) re-estimation.

\input{weeks_coeftable}
\FloatBarrier

% \begin{figure}
% \centering   
%   \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
%   \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
%   \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
% \end{figure}
\FloatBarrier
\clearpage

\subsection*{Gibler (2017)}

Additional information for the Gibler (2017) re-estimation.

\input{gibler_coeftable}
\FloatBarrier

% \begin{figure}
% 	\centering   
% 	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
% 	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
% 	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
% \end{figure}
\FloatBarrier
\clearpage

\section{AME Tutorial}

Using the AMEN function requires formatting data into a particular structure. The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network.

If undirected, the AMEN function has three main inputs:

\begin{itemize}[noitemsep,nolistsep]
    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network. 
    \begin{itemize}
      \item An adjacency matrix describes relationships between nodes in a particular year of data. For example, in an adjacency matrix of interstate MIDs, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had a MID in that year, and '0' otherwise. The diagonal in the adjacency matrix is typically missing. Y is a list of these adjacency matrices for the outcome variable, where each element in the list is a different year of data. 
    \end{itemize}
    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
    \begin{itemize}
      \item An array is a data object in R that contains a series of 'stacked' matrices for each year of data. An array of dimension (2, 3, 4), for example, contains 4 rectangular matrices each with 2 rows and 3 columns. Each matrix in the array describes the relationship between nodes with respect to some covariate. For example, for interstate alliances, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had an alliance in that year, and '0' otherwise. An array contains a matrix of this kind for each covariate going into the model. Xdyad is a list of these arrays, where each element in the list is a different year of data. 
    \end{itemize}
    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
    \begin{itemize}
      \item Each matrix in Xrow has the nodes running down the rows and the covariates in the dataset running along the columns. An entry in the matrix captures the value that a node takes on for each covariate in a given year. For example, if column $j$ is GDP per capita and column $k$ is population, then row $i$, column $j$ measures country $i$'s GDP per capita and row $i$, column $k$ measures country $i$'s population in a given year. Xrow is a list of these matrices, where each element in the list is a different year of data. 
    \end{itemize}
\end{itemize}

If directed, AMEN further requires:

\begin{itemize}[noitemsep,nolistsep]
    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
    \begin{itemize}
      \item Xrow here is nearly identical to the undirected case. The only difference is that rather than each matrix containing covariate data on all nodes, in the directed context each matrix only contains data on the nodes that are acting as senders in that particular year. Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, Xrow would only contain data for those two countries (i.e., Xrow would only have two rows). 
    \end{itemize}
    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
    \begin{itemize}
      \item Xcol here is nearly identical to Xrow, except each matrix contains data on receiver nodes, not sender nodes. Using interstate MIDs as an example, if in 1983 only two countries had MIDs initiated against them, Xcol would only contain data for those two countries (i.e., Xcol would only have two rows). 
    \end{itemize}
\end{itemize}

Beyond the data inputs, the AMEN function requires additional specification:

\begin{itemize}
    \item model: how to model the outcome variable, e.g., `logit'
    \item symmetric: whether the input network is symmetric
    \item intercept: whether to estimate an intercept
    \item nscan: number of iterations of the Markov chain
    \item burn: burn-in period
    \item odens: thinning interval
    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
    \item gof: whether to calculate goodness of fit statistics
\end{itemize}

There is often little theoretical reason to choose a particular value of $\sf{R}$ (above). One strategy is to estimate models at different values of $\sf{R}$ and compare goodness of fit statistics across models.

Given the computational intensity needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis. In addition, providing AMEN function with starting values, either dictated by theory, previous research, or previous runs can also help speed up convergence time.

The code below presents an example of an AME model running in parallel across 4 different levels of $\sf{R}$. Note also that the model is using starting values from a previous run, defined in \textit{startVals0}.

\begin{lstlisting}[language=R]
# running in parallel varying k
imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3

# Run amen in parallel
library(doParallel) ; library(foreach) ; cl=makeCluster(4) ; registerDoParallel(cl)
foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {

  # load previous model run
  load(prevModelFiles[ii])
  # extract start vals
  startVals0 = ameFit$'startVals'
  # dump rest
  rm(ameFit)

  ameFit = ame_repL(
    Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL, 
    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], 
    nscan=imps, seed=1, burn=brn, odens=ods, 
    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,
    periodicSave=TRUE )     
  save(ameFit, file=paste0('model_k', latDims[ii],'_v2.rda') )
}
stopCluster(cl)
\end{lstlisting}