### 2021-01-04 revisionMemo/rev1.tex

\section*{Reviewer 1}

\begin{enumerate}
	\item P1: ``Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases. In practice, such strategies rarely work because they do not directly address the fundamental data generating process.'' The term “rarely work” is too casual.  What problem exactly is motivating this paper?  E.g., the Aronow et al. paper cited at the start of the same paragraph looks at consistent variance estimators for semiparametric regression estimators on dyadic data. By the standards set out by those authors in that paper, the strategies “work” just fine.  I suppose then that the current paper has a different sense of what it means for something “to work.” For example, I am supposing that the current paper is talking about people misconstruing the structural interpretation of a reduced form parameter estimated in a GLM fit to dyadic data.  That would be wholly orthogonal to what Aronow et al. were considering, but then this needs explication.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for taking the time to provide these valuable comments. We have clarified our language in the paper to address this question.  Strategies like those employed by Aronow et al. do not directly model all relevant dependencies in the data but instead adjust the variance estimator. As Aronow et al. explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data'' (p. 16).   While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will lead us to falsely accept the null). The main aims of our research are to account for the relevant dependency structures in the data thereby enabling more accurate inference for covariates and to enable applied scholars using this framework to understand how to better understand the types of dependencies that exist in their data.
		}}
	\end{itemize}
	\item P2: “Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another. This makes sense if each observation is independent of the others.” Revise this opening, since it is perfectly fine to stack observations and then incorporate auxiliary information (such as adjacency matrix) to characterize dependencies between rows.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		 The point is well taken, we have clarified this section (on page 2) to point to the assumption of independence as the primary problem, not the stacking of observations in and of itself.
		}}
	\end{itemize}
	\item P3: “Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.” This is another example of language that is too causal and imprecise, like the opening paragraph. What are “parameter estimates from standard statistical models”?  And “biased” with respect to what target of inference? Again, I think the authors have in mind “bias” with respect to parameters that are meant to have a particular structural interpretation, but this needs to be clarified. Generally, section 2 should be clearer in motivating the problem.  Provide us with a toy example that illustrates the kind of “bias” that the paper has in mind from “standard models”.  For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for this suggestion. We have thoroughly revised the beginning of section two, so that the paper is clearer on the problem that it is seeking out.
	}}
	\end{itemize}
	\item P7: Regarding the specification, this augments a dyadic random effects model, which is quite common, with  the LFM component, which resembles an interactive fixed effects specification (cf. Bai 2009, Ecta) and is based on an SVD factorization of the observed network. It is useful to think of the infeasible estimator and the implications for interpreting the LFM component.  The infeasible estimator takes the transformation f(.), fixed effects parameters, and SRRM random effects as known, in which case the LFM amounts to predictions from a factor model fit to the adjacency matrix of residuals on the scale of the latent outcome.  Would this be a reasonable interpretation? If so it would be useful to draw the connection to factor models commonly in use these days (such as the interactive fixed effects model of Bai, which is the basis of, e.g., generalized synthetic control methods).
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for providing this reference and have added discussion of Bai's work on interactive fixed effects to the manuscript. Within both Bai's work and the model we discuss here the goal is to find ways to use latent factor approaches to model heterogeneity. We'd also note that the recommender system literature from computer science also has its roots in using latent factor models to find low rank approximations. We had noted that already in the manuscript but have tried to make the link clearer in the revised ``Additive and Multiplicative Effect Models for Networks'' section.
		}}
	\end{itemize}
	\item P8: “accounting for this structure is necessary if we are to adequately represent the data generating process.” Continuing with the problems I have with the language above, what does “adequately represent” mean?. P9: “a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.” Okay here is where the paper are starting to clarify goals.  So a) is about causal identification.  But as I understand, the solution here is parametric identification of causal effects.  This means that the issue of misspecification and robustness ought to be taken up.  Similarly, b) is addressed in this paper under a fairly simple parametric data generating process. But again, how robust is the proposed model to situations where the real DGP does not abide by the parametric assumptions (linearity, normally distributed random effects, and a factor model for the residuals of dimension K)? For c), I think most would agree that the “proof” is simply in comparison to whatever other good models are available.  Not however, that in social science applications, forecasting and therefore c) are rarely of interest (at least not in the types of research that appear in political science journals).  As such, from the perspective of what the researchers in discipline mostly cares about, a) and b) are of primary importance.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			For part (a), we have clarified that our study does not address causal identification. We instead focus on providing unbiased estimates of covariates, which improve inference for observational data. We have clarified the goals of the paper in the earlier sections (such as the opening paragraph of page 1) and have also edited P8 to read: ``Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe." }}
		\item \textcolor{blue}{ \emph{
			As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions? We have added an additional simulation section to the appendix that shows how well AME does when parametric assumptions of linearity are violated as discussed in Aronow et al. We also provide simulation evidence to this effect in the end of this section of the revision memo, when responding to another question from the reviewer. }}
		\item \textcolor{blue}{ \emph{
			Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others. }}
	\end{itemize}
	\item P9-10: we need to know exactly how and from what distributions X and W are being drawn to interpret the results here.  Are they independent/orthogonal or no?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for pointing out that the simulation section was unclear. $X$ and $W$ are independent in the simulation that we present in the paper. We have thoroughly revised this section and our explication for how each of the variables were generated in the simulation study. Additionally, we have added in additional sections to the Appendix with simulation setups where $X$ and $W$ are related in alternative ways. These additonal sections come in response to requests from both Reviewer 1 and 2, and we provide more detail on them later in the memo.
		}}
	\end{itemize}
	\item Section 4 needs to incorporate an assessment of robustness to misspecification.  See the Aronow et al. paper for an example of how to do this, where they study the performance of a dyadic random effects model.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have added a simulation section to the Appendix in which we have incorporated an additional assessment of robustness to misspecification such as the one considered in Aronow et al. Specifically, Aronow et al note that ``fitting a linear approximation to mildly non-linear data'' can be problematic for dyadic random effect models. We agree on the importance of testing this in the context of our simulation. Aronow et al incorporate non-linear misspecification by adding in squared version of the key dyadic homophily variable to the DGP for their simulations: $Y_{i,j} = \beta_{0} + \beta_{1} X_{i,j} + \beta_{2} X_{i,j}^{2} + \epsilon_{i,j}$. Here $X$ is their representation of homophily and is considered observed, while $X^{2}$ is their check for robustness to misspecification and is unobserved. We incorporate the same type of misspecification into our probit simulation setup: $Z_{i,j} =  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}$, where $W$ now instead of being drawn from a separate normal distribution is just the squared version of $X$. For the simulation we set values of $\mu$, $\beta$, and $\gamma$ to -2, 1, and 1, respectively. Results are shown below in Figures~\ref{fig:ameBias_asa} and \ref{fig:ameCalib_asa} below. As in the paper, along the x-axis we implement a standard model that does not take any steps to account for dependencies, AME, and the Oracle model, and as in the paper we find that the AME approach provides a number of benefits in terms of bias and coverage over the standard approach of not dealing with dependencies.
	}}

	\begin{figure}[ht]
		\centering
		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
		\label{fig:ameBias_asa}
		\includegraphics[width=1\textwidth]{ameSimBias_all_asaProbit.pdf} \\
	\end{figure}

	\begin{figure}[ht]
		\centering
		\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
		\label{fig:ameCalib_asa}
		\includegraphics[width=1\textwidth]{ameSimCover_all_asaProbit.pdf} \\
	\end{figure}

	\end{itemize}
\end{enumerate}

\FloatBarrier
\clearpage