### 2021-01-04 revisionMemo/rev2.tex

\section*{Reviewer 2}

\begin{enumerate}
	\item The additive and multiplicative effects (AME) model presented in the paper allows for a better treatment of first, second and third order dependencies in the data, and account for unobserved sources of bias in the data. The presentation of the AME estimator is clear, and the simulations provide clear evidence that the model outperforms naive OLS/GLM models fitted to the same data. The paper will be of great interest to the PSRM audience and is likely to have a strong impact on future research in international relations. I do, however, have a few suggestions for revisions that will hopefully strengthen the paper’s contributions.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for their comments and suggestions.
		}}
	\end{itemize}
	\item The paper shows that the AME model performs well in comparison to the “oracle” model where all sources of dependencies in the data are accounted for, which is reassuring. Yet this comparison feels incomplete. First, the naive model is clearly misspecified given the underlying data generating process. A large number of dependencies in international relations are linked to geography. What would happen if the dependencies were modeled using spatial econometric techniques?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			The reviewer raises an important connection to the spatial literature. The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria. With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l). In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (Plumper and Neumayer (2010) provide an in-depth discussion on varying operationalizations of spatial lags). If there is a particular type of spatial dependence that the reviewer would like us to simulate, however, we are happy to do so.
			}}
	\end{itemize}
	\item It would help if the author discussed in more depth the properties of the AME estimator in the presence of unobserved dependencies. One would imagine that the oracle model, where all the unobserved factors and dependencies are accounted for should always outperform a parametric or semi-parametric estimator. Relatedly, it also would help to have a more technical discussion of the properties of AME models, particularly in terms of the bias-variance trade off.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			Estimation certainly gets more challenging when the omitted variable is correlated with the observed variable. In the case of normal linear regression, we can show that in general a regression coefficient is unbiased for its ``unconditional effect'', which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with. For example, suppose in our simulation setup $X_{i,j}$ is correlated with $W_{i,j}$ via the relation $W_{i,j} = \alpha X_{i,j} + e_{i,j}$. Then: }}

			\begin{align*}
			  Y_{i,j} & = \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j} \\
			  &= \mu + (\gamma \alpha+\beta ) X_{i,j} + (\gamma Z_{i,j} + \epsilon_{i,j})  \\
			  &= \mu + \tilde \beta X_{i,j} + \tilde \epsilon_{i,j}
			\end{align*}

		\item \textcolor{blue}{ \emph{
			The naive model is in some sense still ``correct'' as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $X$ on $Y$. However, there is no way to eliminate this extra bias without making further assumptions about the nature of the omitted variable. Additonally, in the binary case with the probit setup, other sources of bias get introduced as well because of the nonlinearity of the population moments as a function of the parameters. An interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors. An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.}}

		\item \textcolor{blue}{ \emph{
			For now though to show the amount of bias that gets introduced into our setup when $X$ and $W$ are correlated we repeat our simulation exercise but induce varying levels of correlation between $X$ and $W$. Results with $X$ and $W$ correlated at 0.4 and then 0.7 are shown below, respectively.}}

			\begin{figure}[ht]
			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.4 in this case.}
			\label{fig:ameBias_corrMed}
			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitMed.pdf} \\
			\end{figure}

			\begin{figure}[ht]
			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.7 in this case.}
			\label{fig:ameBias_corrHi}
			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitHi.pdf} \\
			\end{figure}
			\FloatBarrier

		\item \textcolor{blue}{ \emph{
			In terms of the bias-variance tradeoff discussion, this enters most prominently as a consideration for AME when choosing the dimensions of the latent factor space, specifically $K$ in the model. Everytime we add another dimension to the latent factor space the number of parameters in the model increases notably. For example, in a directed network with a choice of $K=1$ there are $2 \times n$ parameters introduced from the LFM portion of the model. With $K=2$ this number increases by another $2 \times n$. Increasing K certainly comes with a cost in terms of the number of extra parameters that are introduced into the model.
		}}
	\end{itemize}
	\item The author could also underscore a related issue, which usually escapes empirical IR scholars: the importance of developing theories that account for the direct and indirect effects of interactions in a setting where the units are interconnected, and where the effects on one unit have the potential to impact other units in the system. A traditional approach is to throw as many observable variables as possible as controls, without much consideration on where they fit in the model. As the comparison of the AME and oracle models suggests, a careful modeling of spillovers and dependencies can result in better estimates of the underlying relationships.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We agree with this suggestion. We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p. 13), as well as in the conclusion of our paper. We agree that innovations in dependence modeling can and should speak to theoretical advancements that, like our model, take into account the important relational and interdependent nature of political processes.
		}}
	\end{itemize}
	\item The paper emphasizes the importance of estimating the random effects for theory building and hypothesis test. A longer elaboration and illustration of this point would be very helpful.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have added an additional interpretation of the random effects on pg. 19 in our discussion of Reiter and Stam. (``Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict. Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest. In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for Colgan's (2010) theory that revolutionary petrostates are more aggressive.")
		}}
	\end{itemize}
	\item I would like to see a more detailed discussion of the setup and choices made for the Monte Carlo simulations. What type of network do the simulated data model? How does the network in the simulation relate to typical networks in IR studies, and to the ones in the replicated papers.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for this comment and have revised our simulation setup as a result. Most IR studies into networks have been interested in modeling binary outcomes (e.g., indicators for international agreements between countries or conflicts between countries), thus we we have revised our simulation to have a binary dependent variable. Key results remain the same.
		}}
		\item \textcolor{blue}{ \emph{
		Additionally, we have thoroughly revised our description to make it clearer how each of the variables in the simulations were generated.
		}}
	\end{itemize}
	\item The criteria for choosing the studies for replication is unclear and needs a better justification. I would have liked to see a larger set of replications particularly in areas and topics where the data generating process, the dependencies among units varies, or the relationship between the units is hierarchical, cross-classified or multilevel.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We selected our cases based on a few criteria. We focus on studies that are explicitly about International Relations, were published since the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we also decided to focus on one journal, the APSR). We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience. In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam 2003), third-order dependencies (Weeks 2012), precision of fixed effects estimates (Gibler 2017).  While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, we believe that they are beyond the scope of this paper at this time.
		}}
	\end{itemize}
	\item Given the importance of promoting the use of the AME method I would encourage the author to revise the tutorial presented in Appendix C to make it more accessible to a larger audience.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		The reviewer raises a valid point. We have made a number of revisions and extensions to the tutorial at the end of the Appendix so that applied scholars can better take advantage of AME.
		}}
	\end{itemize}
\end{enumerate}