### 2021-01-04 revisionMemo2/rev2.tex
revision memo for second round

\section*{Reviewer 2}

\begin{enumerate}
	\item I would have liked to see a justification for the choice of papers used in the replications. It is still unclear. Are these cases where the theory fail to characterize the direct and indirect effects, and implicitly assume that the latent dependencies are nuisances?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		something, something
		}}
	\end{itemize}
	\item The presentation of the AME model clearly states that the Xs are exogenous; it is likely that they are not in replications. Still, the AME estimates (which in the simulations do not seem to outperform the standard method for estimating the betas when correlation between X and W is high) suggest that the models in the original papers are misspecified, and that these differences are likely driven by a better modeling of the DGP. This is a very important take-home point.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			something, something
			}}
	\end{itemize}
\end{enumerate}